Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality

### 2015-06-16 content/contribution/continuous-datalog.tex
concluded section on continuous datalog by bringing up the concept of "valid windows" as a generalisation of "sliding windows" (i.e. windows of data valid by some definition as opposed to windows of data valid by the specific definition of a duration from receipt).

	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
	\begin{description}
		\item[Continuous Datalog Program\label{def:continuous datalog: CDP}]\hfill\\
			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${t \in \mathbb{N}}$ as arguments.
			\begin{equation*}
				CDP = \{A,S,R\} \vDash \{A',P,N\}
			\end{equation*}
		\item[State of a ${CDP}$\label{def:continuous datalog: CDPt}]\hfill\\
			The state of a ${CDP}$ at a given instant ${t \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
			As both $A$ and $W^{S,R}_{i}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $t$ by $T^{CD}$.
			The instantaneous entailments of a $CDP_{i}$ are the union of the persistent entailments of ${A \vDash A'}$ (which are guaranteed to be entailed by every state, as Positive Datalog is monotonic) and the set $E_{i}$ of instances of the transient entailments that are justified at least in part by some subset of $W^{S,R}_{i}$. 
			\begin{equation*}
				T^{CD} \left( CDP, t \right) = T^{D} \left( CDP_{i} \right)
			\end{equation*}
			\begin{equation*}
				CDP_{i} = A \cup W^{S,R}_{i} \vDash A' \cup E_{i}
			\end{equation*}
	\end{description}
	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.
	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: positive window increment}
			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $t$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${t-1}$.
			\begin{equation*}
				W^{P,1}_{i} = \left( A' \cup E_{i} \right) \setminus \left( A' \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
			\end{equation*}
		\item\label{axiom:continuous datalog: negative window increment}
			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at instant $t$.
			\begin{equation*}
				W^{N,1}_{i} = \left( A' \cup E_{i-1} \right) \setminus \left( A' \cup E_{i} \right) = E_{i-1} \setminus E_{i}
			\end{equation*}
		\item\label{axiom:continuous datalog: entailment precedes negation}
			Facts may only be negated after they have been entailed.
			As such, a window over the stream of entailments $P$ at any given instant ${t \in \mathbb{N}}$ will be disjoint from the 
			\begin{equation*}
				\forall Y \geq 1 \left( W^{P,1}_{i} \cap W^{N,1}_{i-Y} = \varnothing \right)
			\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

	From this, we can see that each instance $e$ of an entailment in some $E_{i}$ may take the same form as any other fact in a stream of facts, being ${e = p(s,o):t}$, where $t$ is the real time associated with the instant in which $e$ is first entailed by the ${CDP}$ (i.e. it appears in $E_{i}$ but not $E_{i-1}$, and so is part of $W^{P,1}_{i}$).
	As such, the stream of entailments $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time of the entailments but by the time at which they \emph{cease} to be entailed (i.e. they appear in $E_{i-1}$ but not $E_{i}$, and so is part of $W^{N,1}_{i}$).

	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
	However, it can be shown from these axioms that it is also possible to reconstruct the instantaneous set of entailments of a ${CDP}$ at a given instant from the contents of windows of sufficient range over its entailed streams $P$ and $N$ at that instant.

	\subimport{continuous-datalog/}{proof-of-entailment-semantics.tex}
\end{nestedsection}

>>>>

	Given these definitions, a Continuous Datalog program may be formally defined, both in terms of streams and in terms of its instantaneous states:
	\begin{description}
		\item[Continuous Datalog Program\label{def:continuous datalog: CDP}]\hfill\\
			A Continuous Datalog Program (${CDP}$) is composed of a set of Positive Datalog axioms $A$, a set of streams of facts $S$ and the range ${R \in \mathbb{N}}$ of the windows over each stream in $S$.
			This specifies that the streamed data from $S$ is to be considered valid from its arrival for the duration represented by $R$.
			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.
			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${t \in \mathbb{N}}$ as arguments.
			\begin{equation*}
				CDP = \{A,S,R\} \vDash \{A',P,N\}
			\end{equation*}
		\item[State of a ${CDP}$\label{def:continuous datalog: CDPt}]\hfill\\
			The state of a ${CDP}$ at a given instant ${t \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
			As both $A$ and $W^{S,R}_{i}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $t$ by $T^{CD}$.
			The instantaneous entailments of a $CDP_{i}$ are the union of the persistent entailments of ${A \vDash A'}$ (which are guaranteed to be entailed by every state, as Positive Datalog is monotonic) and the set $E_{i}$ of instances of the transient entailments that are justified at least in part by some subset of $W^{S,R}_{i}$. 
			\begin{equation*}
				T^{CD} \left( CDP, t \right) = T^{D} \left( CDP_{i} \right)
			\end{equation*}
			\begin{equation*}
				CDP_{i} = A \cup W^{S,R}_{i} \vDash A' \cup E_{i}
			\end{equation*}
	\end{description}
	The following axioms define the semantics of the streams of entailments $P$ and and negations $N$ of a ${CDP}$ in terms of the instantaneous entailments of the sequence of states of that continuous program.
	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: positive window increment}
			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $t$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${t-1}$.
			\begin{equation*}
				W^{P,1}_{i} = \left( A' \cup E_{i} \right) \setminus \left( A' \cup E_{i-1} \right) = E_{i} \setminus E_{i-1}
			\end{equation*}
		\item\label{axiom:continuous datalog: negative window increment}
			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at instant $t$.
			\begin{equation*}
				W^{N,1}_{i} = \left( A' \cup E_{i-1} \right) \setminus \left( A' \cup E_{i} \right) = E_{i-1} \setminus E_{i}
			\end{equation*}
		\item\label{axiom:continuous datalog: entailment precedes negation}
			Facts may only be negated after they have been entailed.
			As such, a window over the stream of entailments $P$ at any given instant ${t \in \mathbb{N}}$ will be disjoint from the 
			\begin{equation*}
				\forall Y \geq 1 \left( W^{P,1}_{i} \cap W^{N,1}_{i-Y} = \varnothing \right)
			\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

	From this, we can see that each instance $e$ of an entailment in some $E_{i}$ may take the same form as any other fact in a stream of facts, being ${e = p(s,o):t}$, where $t$ is the real time associated with the instant in which $e$ is first entailed by the ${CDP}$ (i.e. it appears in $E_{i}$ but not $E_{i-1}$, and so is part of $W^{P,1}_{i}$).
	As such, the stream of entailments $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time of the entailments but by the time at which they \emph{cease} to be entailed (i.e. they appear in $E_{i-1}$ but not $E_{i}$, and so are part of $W^{N,1}_{i}$).

	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
	However, it is proven in \refapp{semantics: proof for entailments} that, from these axioms, it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]
	The minimum range $X$ is such to achieve the \emph{complete} set of entailment instances $E_{i}$ of the ${CDP}_{i}$.
	In addition, the minimum range $Y$ is such to achieve the \emph{correct} set $E_{i}$.
	As such, a \emph{valid window} that contains only and exactly those instances of entailments in $E_{i}$ at every instant $i$ may be trivially maintained from the streams $P$ and $N$ continuously entailed by some ${CDP}$.
\end{nestedsection}
